### 2018-03-28 response.tex

\setlength{\paperwidth}{210mm}
\setlength{\paperheight}{297mm}% fixed.

\documentclass{article}

\usepackage[review]{myreviewpckg}

\usepackage{textcomp}
\usepackage{simplemargins}

\clubpenalty=10000  % Orphan - First of paragraph left behind
\widowpenalty=10000 % Widow  - Last of paragraph sent ahead

\setallmargins{1in}

\begin{document}

% ******************************************************************
% *************************** REVIEWER 1 ***************************
% ******************************************************************

\begin{center}
	\bf
	\large
	Authors' Response to Reviewer 1
\end{center}

\noindent
We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

\introcomment{~}{%
This is an interesting article for the readers of the BSSA journal. The authors propose a new approach based on machine learning techniques to determine the importance of validation metrics for ground motion simulation. The field of ground motion simulation validation (GMSV) is relatively new and of interest to both seismologists who develop simulation models and engineers who would like to use these simulations in engineering applications. One of the most challenging questions to address is "which validation metric(s) is best to use?" as there are so many different metrics in this field and so many different opinions. It has been difficult to address this question because the answer varies depending on the application of the simulations (e.g., GMPE development, PSHA, structural analysis response of interest, etc). This paper is interesting because it provides a statistical approach to reduce the number of validation metrics and select a few representatives from a bigger set without compromising the final validation results. The authors have done a great job providing background on the topic. However, their application is very narrow, they only consider the 11 Anderson2004+ parameters (as they've noted there are other parameters) for one simulation method for one earthquake, and as a result their conclusions are only valid for this particular case. They have mentioned this limitation in the conclusions, but the abstract and introduction do not explain the limitations and the application-dependency of the proposed approach. Both Abstract and Introduction need modification to specify the application for which the results and conclusions are true (i.e., the
application here is to narrow down the 11 Anderson2004+ parameters to select fewer parameters that give similar results as the original 11 for the simulations they chose in this paper). Currently, the Abstract and Introduction give the impression that the conclusions are true for any engineering application or simulation method. Furthermore, there should be a short discussion in the paper on how the users can apply the same method for different applications, for example for structural responses such as drift ratio or responses that are sensitive to high-frequency content of simulations (theoretically, this method should be applicable to any response of interest if they replace the GOF scores used in this study? If true this should be mentioned as it will be of interest to others).
}

\response{%
Will be added. 
}

\comment{1}{%
The terminologies used in the paper are not clear in many cases. For example, what are "features" "attributes" "dimensions" "trees, nodes and leafs" "class"? It was very difficult to
read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper. I was confused what
"data" and "results" were referring to: GOF score data? GOF metric data? GOF category? Something other than GOF? I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
}

\response{%
response
}

\comment{2}{%
The terms "common practice" or "common standards" have been mentioned several times. Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree
with some statements but I agree that others may), I think these terms should be removed. Just say this is your opinion, or it is based on the references that you have listed.
}

\response{%
Response 
}

\comment{3}{%
In abstract, "semi-supervised" and "supervised" learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
and what does that mean?
}

\response{%
Response
}

\comment{4}{%
In introduction, the difference between "verification" and "validation" should be discussed.
}

\response{%
Response
}

\comment{5}{%
In abstract and introduction, what does a "direct application" mean? it's not clear. Similarly, in Validation Metrics section: "direct quantitative comparisons" and "indirect statistical analysis"
}

\response{%
Response
}

\comment{6}{%
Page 4, line 53: add "relatively" before "well understood"
}

\response{%
Response
}

\comment{7}{%
The two questions posed in the introduction lines 59-61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e.,
has proposed them). The second question is the main question and the one being addressed in this paper.
}

\response{%
Response
}

\comment{8}{%
You are treating these metrics as "independent variables", but as you have also noted many of them are not. For example, Arias intensity and energy are correlate. Does this compromise your conclusions? If not explain why in the Results section.
}

\response{%
Response
}

\comment{9}{%
Page 7, line 133: delete the last sentence
}

\response{%
Response
}

\comment{10}{%
Page 14, line 310: you say "... the latter can essentially be discarded..." I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are
independent of each other (one gives information that the other one doesn't)?
}

\response{%
Response
}

\comment{11}{%
Page 14 line 323: It's not clear how you "replicate data", provide more detail.
}

\response{%
Response
}

\comment{12}{%
I don't think "confusion matrix" given in tables is explained in the paper!
}

\response{%
Response
}

\comment{13}{%
Page 17, lines 391-394: this sentence should be re-worded, what is an "accurate conclusion"? You should be very specific that your conclusion is that instead of using all 11 validation metrics you can select a few in order to get the same categorization as all 11 metrics would have given you.}

\response{%
Response
}

\comment{14}{%
Page 18, line 405: this is implying that fig 10 is similar to fig2! I personally cannot see this in the figures! They look very different. It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
}

\response{%
Response
}

\comment{15}{%
Figure 8: I don't know what you mean by y-label "decision attributes". Do you mean "the number of GOF metrics"? why there is no data point for 1 or 11?
}

\response{%
Response
}

\comment{16}{%
The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters
describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from
response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa
being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your
conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
}

\response{%
Response
}

Typos:\\
\\

\comment{1}{%
Page 2, line 8: change focused to focus
}

\response{%
Response
}

\comment{2}{%
Page 2, line 20: change classifiers to classifier
}

\response{%
Response
}

\comment{3}{%
Page 13, line 185: change measure to measured
}

\response{%
Response
}

\comment{4}{%
Page 16, line 354: the last C5 should be C6?
}

\response{%
Response
}

% ******************************************************************
% *************************** REVIEWER 2 ***************************
% ******************************************************************

\newpage

\begin{center}
	\bf
	\large
	Authors' Response to Reviewer 2
	\end{center}

\noindent
We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

\introcomment{~}{%
Review of Khoshnevis and Taborda for BSSA
}

\introcomment{~}{%
Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak
ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear. This paper uses machine learning techniques and eleven possible metrics for comparing (recorded and simulated) ground motions to determine the parameters that are most predictive of goodness-of-fit. This is an interesting paper, and will help future researchers
prioritize which metrics to focus on when validating ground motions simulations. The paper is generally well written. As a strong ground motion seismologist (i.e., I assume I'm part of the audience you hope to reach), there is a lot of machine learning-related jargon that makes certain sections difficult to follow. It would also be useful to have some discussion regarding which parameters were found to be repetitive, incase future authors wish to look at more than the four recommended parameters.
}

\comment{1}{%
L. 8 "... could \textbf{focus} ..."
}

\response{
Response
}

\comment{2}{%
L. 11 "... the objective \textbf{of} ..."
}

\response{
Response
}

\comment{3}{%
L. 25-41 The frequent use of phrases such as "Among these methods,"; "Among the latter,"; "In this category,"; and "Within this group," make the text difficult to follow.
}

\response{
Response
}

\comment{4}{%
L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.
}

\response{
Response
}

\comment{5}{%
L. 59-61/L. 391-394 It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
}

\response{
Response
}

\comment{6}{%
L. 61 This sentence should end with a "?"
}

\response{
Response
}

\comment{7}{%
L. 98-99 It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum). Do you follow the guidelines provided by Anderson? If
so, what frequency range/bands are you examining?
}

\response{
Response
}

\comment{8}{%
L. 128/132/137/etc. "... the three component\textbf{s} ..."
}

\response{
Response
}

\comment{9}{%
L. 236 Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is L. 328.
}

\response{
Response
}

\comment{10}{%
L. 236 Remove "A matter we discuss later."
}

\response{
Response
}

\comment{11}{%
L. 241-243 It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
}

\response{
Response
}

\comment{12}{%
L. 257-258 I don't understand this sentence. What is a continuous attribute?
}

\response{
Response
}

\comment{13}{%
L. 259-264 This paragraph contains a lot of jargon.
}

\response{
Response
}

\comment{14}{%
L. 300 Isn't this also true for the C5-C7 combination?
}

\response{
Response
}

\comment{15}{%
L. 328 What is "a strong pruning process"?
}

\response{
Response
}

\comment{16}{%
L. 433 "We present the result\textbf{s} of..."
}

\response{
Response
}

\comment{17}{%
L. 435 "... goal of prioritiz\textbf{ing} and reduc\textbf{ing}..."
}

\response{
Response
}

\comment{18}{%
L. 443 Remove "Similarly" or "similar" (redundant)
}

\response{
Response
}

\comment{19}{%
L. 450 "... remain valid."
}

\response{
Response
}

\comment{20}{%
L. 455 "... as the decisive \textbf{parameters} ..." ? (This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., "This latter point", "One the one hand",
"on the other hand."
}

\response{
Response
}

\end{document}

### 2018-03-30 response.tex

\clubpenalty=10000  % Orphan - First of paragraph left behind
\widowpenalty=10000 % Widow  - Last of paragraph sent ahead

\setallmargins{1in}

\begin{document}

% *****************************************************************
% *************************Associate Editor *********************
% *****************************************************************

\begin{center}
	\bf
	\large
	Authors' Response to Associate Editor (Dr. Thomas Brocher)
\end{center}

\noindent
We thank Dr. Thomas Brocher for his comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

\introcomment{~}{%
I agree with the peer reviewers that this paper addresses a topical subject of great importance to BSSA readers. The science is in the process of transitioning from purely observational based analysis of ground motions to a mixture of observations and simulations of ground motions. The paper is a useful contribution to this discussion, is generally well written, and carefully
describes the methodology used and its limitations. The peer reviewers have made some very helpful suggestions for improving the paper that I hope the authors will strongly consider.\\
In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions. In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit. It clearly shows that only these metrics are change significantly with the goodness of fit (P,F, G, E). Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this
relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic. In fact, in my opinion, Figure 6 is a simple and easy way of determining which metrics are diagnostic that might even eliminate the need for the decision trees. In any case, more
discussion of this figure might help the authors make the case for the main conclusions.
}

\response{%

Thanks for the constructive suggestions. We agree with you and peer reviewers that the technical terms are not commonly used in seismology and strong motion earthquake engineering. The confusion happened because of integrating three different topics into a method to carry out this research. One technical term in one field is different in another field and that cause a confusion. For example cluster, class, and label all are referring to the same concept, however, it depends on the application. We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
We also modified Figure 6 in order to easily track the variation of different metrics for final classes. We do agree with you that Figure 6 clearly gives a good understanding about the metrics, however, we add decision trees, because they can convert the qualitative judgment to quantitative conditional form to be able to exactly decide on the final class based on their numerical GOF scores. \\
We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.\\
}

% ******************************************************************
% *************************** REVIEWER 1 ***************************
% ******************************************************************
\newpage
\begin{center}
	\bf
	\large
	Authors' Response to Reviewer 1
\end{center}

\noindent
We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

@@

\response{%
The reviewer pointed out a very important topic that we believe we did not discuss it well in the manuscript. The manuscript presents steps of moving from considerable amount of data with specialties knowledge (background knowledge) to a final product that gives a better understanding about the data. According the reviewer suggestion we added a paragraph  to discuss this important point. Also we modified abstract an introduction for acknowledging the limitations with the study. 
}

@@

\comment{1}{%
The terminologies used in the paper are not clear in many cases. For example, what are "features" "attributes" "dimensions" "trees, nodes and leafs" "class"? It was very difficult to
read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper. I was confused what
"data" and "results" were referring to: GOF score data? GOF metric data? GOF category? Something other than GOF? I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
}

\response{%
We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering. The confusion happened because of integrating three different topics into a method to carry out this research. One technical term in one field is different in another field and that cause a confusion. For example cluster, class, and label all are referring to the same concept, however, it depends on the application. We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
}

@@

\comment{2}{%
The terms "common practice" or "common standards" have been mentioned several times. Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree
with some statements but I agree that others may), I think these terms should be removed. Just say this is your opinion, or it is based on the references that you have listed.
}

\response{%
Thanks for the comment. We fixed it. 
}

\comment{3}{%
In abstract, "semi-supervised" and "supervised" learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
and what does that mean?
}

\response{%
We briefly explained these terms \textbf{Data Analysis Method} section. We also added clarification about semi-supervised and supervised learning into  \textbf{Introduction} section.
}

@@

\comment{4}{%
In introduction, the difference between "verification" and "validation" should be discussed.
}

\response{%
A discussion about verification and validation is added. 
}

@@

\comment{5}{%
In abstract and introduction, what does a "direct application" mean? it's not clear. Similarly, in Validation Metrics section: "direct quantitative comparisons" and "indirect statistical analysis"
}

\response{%
\color{red}{Ricardo, please address this comment.}
}

@@

\comment{6}{%
Page 4, line 53: add "relatively" before "well understood"
}

\response{%
Added. Thanks.
}

\comment{7}{%
The two questions posed in the introduction lines 59-61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e.,
has proposed them). The second question is the main question and the one being addressed in this paper.
}

\response{%
We modified the text for clarification and also to acknowledge each individual proposed metrics' importance. 
}

\comment{8}{%
You are treating these metrics as "independent variables", but as you have also noted many of them are not. For example, Arias intensity and energy are correlated. Does this compromise your conclusions? If not explain why in the Results section.
}

\response{%
This is a very interesting comment. You are totally right. There are fairly high correlation between Arias intensity and energy integral. However, decision tree algorithm is robust enough to realize this similarity and pick the best one in each decision making process. As we remove the peak ground acceleration and response spectra from the prediction model, the algorithm looks for other best parameters. In case of hundreds of metrics it would be a reasonable choice to do feature selection analysis and choose one metric out of several correlated metrics. However, in this study, we only deal with 11 metrics and the results show that it works fairly well. 
}

@@

\comment{9}{%
Page 7, line 133: delete the last sentence
}

\response{%
Deleted. Thanks.
}

\comment{10}{%
Page 14, line 310: you say "... the latter can essentially be discarded..." I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are
independent of each other (one gives information that the other one doesn't)?
}

\response{%
In that specific subspace (C1,C7), C1 does not help in clustering process. That is why we call it irrelevant feature, because one can define the clusters based on only defining boundaries on C7. C1 can be important, however, it does not add any positive or negative values in decision making process in this subspace analysis. 
}

@@

\comment{11}{%
Page 14 line 323: It's not clear how you "replicate data", provide more detail.
}

\response{%
Thanks. More details provided. 
}

\comment{12}{%
I don't think "confusion matrix" given in tables is explained in the paper!
}

\response{%
In \textbf{Decision Trees} section we provided a definition of confusion matrix for the simplest case of two categories. The idea can be extended for more than two (in our case four) categories. 
}

@@

\comment{13}{%
Page 17, lines 391-394: this sentence should be re-worded, what is an "accurate conclusion"? You should be very specific that your conclusion is that instead of using all 11 validation metrics you can select a few in order to get the same categorization as all 11 metrics would have given you.}

\response{%
You are right. It could be misleading. We re-worded the sentence. 
}

@@

\comment{14}{%
Page 18, line 405: this is implying that fig 10 is similar to fig2! I personally cannot see this in the figures! They look very different. It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
}

\response{%
That is correct. The statement convey part of the message. Although we expect to see similarities between the current results and previous studies, we keep in mind that this study is based on lack of consensus on using GOF metrics for validation purposes. Since the previous study is representing a uniform average of all metrics, we do not expect to match all stations. Therefore, we prefer to keep it as numerical. We modified the text accordingly. 
}

@@

\comment{15}{%
Figure 8: I don't know what you mean by y-label "decision attributes". Do you mean "the number of GOF metrics"? why there is no data point for 1 or 11?
}

\response{%
That is right. It should be the number of decision attributes. We modified the figure accordingly. 
}

@@

\comment{16}{%
The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters
describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from
response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa
being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your
conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
}

\response{%
That is a very good suggestion. We excluded C8(response spectra) and C5(peak acceleration) from analysis and reproduce similar trees with same complexities in terms of pruning process with previous four trees. As we expected, most important, metrics are still C4 and C6.  C11 also become somewhat effective in the most complicated tree. 
}

@@

% ******************************************************************
% *************************** REVIEWER 2 ***************************
% ******************************************************************

\newpage

\begin{center}
	\bf
	\large
	Authors' Response to Reviewer 2
	\end{center}

\noindent
We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

@@

\introcomment{~}{%
Review of Khoshnevis and Taborda for BSSA
}

\introcomment{~}{%
Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak
ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear. This paper uses machine learning techniques and eleven possible metrics for comparing (recorded and simulated) ground motions to determine the parameters that are most predictive of goodness-of-fit. This is an interesting paper, and will help future researchers
prioritize which metrics to focus on when validating ground motions simulations. The paper is generally well written. As a strong ground motion seismologist (i.e., I assume I'm part of the audience you hope to reach), there is a lot of machine learning-related jargon that makes certain sections difficult to follow. It would also be useful to have some discussion regarding which parameters were found to be repetitive, incase future authors wish to look at more than the four recommended parameters.
}

\response{
We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering. The confusion happened because of integrating three different topics into a method to carry out this research. One technical term in one field is different in another field and that cause a confusion. For example cluster, class, and label all are referring to the same concept, however, it depends on the application. We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
We also studied the models without response spectra (C8) and without response spectra (C8) and peak ground acceleration (C5).\\ 
}

@@

\comment{2}{%
L. 11 "... the objective \textbf{of} ..."
}

\response{
Added. Thanks.
}

\comment{3}{%
L. 25-41 The frequent use of phrases such as "Among these methods,"; "Among the latter,"; "In this category,"; and "Within this group," make the text difficult to follow.
}

\response{
\color{red}{Ricardo, please address this comment.}
}

\comment{4}{%
L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.}

\response{
Thanks for the suggestion. We moved the description of the Anderson (2004) method (i.e., L. 48+) to the mentioned location. It reads better now. 
}

@@

\comment{5}{%
L. 59-61/L. 391-394 It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
}

\response{
Thanks for pointing this out. Since Anderson developed metrics for engineering application, we assumed it is implicitly mentioned. However, we agree that it should be explicitly clarified. We modified the text accordingly. 
}

@@

\comment{6}{%
L. 61 This sentence should end with a "?"
}

\response{
Fixed. Thanks.
}

\comment{7}{%
L. 98-99 It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum). Do you follow the guidelines provided by Anderson? If
so, what frequency range/bands are you examining?
}

\response{
In this study we only use broadband results. We added a clarification into the text. Thanks for the comment.  
}

@@

\comment{11}{%
L. 241-243 It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
}

\response{
We added a brief description about them. Thanks.
}

\comment{12}{%
L. 257-258 I don't understand this sentence. What is a continuous attribute?
}

\response{
We meant to say numerical continues attribute. Thanks for the comment. Fixed. 
}

@@

\comment{13}{%
L. 259-264 This paragraph contains a lot of jargon.
}

\response{
Thanks for the comment. Additional clarification added. 
}

@@

\comment{14}{%
L. 300 Isn't this also true for the C5-C7 combination?
}

\response{
That is right. Added.
}

\comment{15}{%
L. 328 What is "a strong pruning process"?
}

\response{
We modified it to heavy pruning process and also added its explanation. 
}

@@

\comment{19}{%
L. 450 "... remain valid."
}

\response{
Extra "s" deleted. Thanks.
}

\comment{20}{%
L. 455 "... as the decisive \textbf{parameters} ..." ? (This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., "This latter point", "One the one hand",
"on the other hand."
}

\response{
\color{red}{Ricardo, please address this comment.}
}

### 2018-03-30 abstract.tex
First draft of revision.

% 
It has become common practice to validate ground motion simulations based on a variety of time and frequency metrics scaled to quantify the level of agreement between synthetics and data, or with respect to reference solutions. There is, however, no agreement about the importance or weight that it ought to be given to each metric. This leads to their selection often being subjective, either based on intended applications or personal preferences. As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could focused on a reduced number of metrics. We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation. In particular, we study the relationships that exist between eleven different metrics. In the process carried out, these metrics are understood as part of a multi-dimensional space. We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects. This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following common practice. We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories. These decision trees can be understood as rapid predictors of the quality of a simulation, or as a data-informed classifiers that can help prioritize validation metrics. Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.

>>>>

% 
It has become common practice to validate ground motion simulations based on a variety of time and frequency metrics scaled to quantify the level of agreement between synthetics and data, or with respect to reference solutions. There is, however, no agreement about the importance or weight that it ought to be given to each metric. This leads to their selection often being subjective, either based on intended applications or personal preferences. As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \myremove{focused} \myrevision{focus} on a reduced number of metrics. We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective \myrevision{of} prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation. \myrevision{We use a dataset of three physic-based regional scale ground motion simulations based on one earthquake and three velocity models with eleven goodness of fit metrics.} In particular, we study the relationships that exist between \myremove{eleven} \myrevision{these} different metrics. In the process carried out, these metrics are understood as part of a multi-dimensional space. We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects. This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following \myrevision{previous studies} \myremove{common practice}. We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories. These decision trees can be understood as rapid predictors of the quality of a simulation, or as  \myremove{a} data-informed classifiers that can help prioritize validation metrics. Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration. \myrevision{Adding other simulation datasets (i.e., from different earthquakes, regions, models, metrics, etcetera) will improve the results in terms of developing a sufficiently sound and robust decision tree, however, the procedural steps laid out in this study, nonetheless, remain valid.}

### 2018-03-30 conclusions.tex
First draft of revision.

We present the result of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritize and reduce the number of metrics, and develop an application independent decision algorithm. As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset. In particular, the proposed algorithm uses three (disjunctive decision) steps based on the values obtained for the metrics of the total energy and acceleration response spectra. we also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies. Similarly, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.

We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake in a particular region. In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, etcetera) in order to arrive to a sufficiently sound and robust decision tree. The procedural steps laid out here, nonetheless, remains valid. We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.

>>>>

We present the result\myrevision{s} of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of \myremove{prioritize} \myrevision{prioritizing} and \myremove{reduce} \myrevision{reducing} the number of metrics, and develop an application independent decision algorithm. As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset. In particular, the proposed algorithm uses three (disjunctive decision) steps based on the values obtained for the metrics of the total energy and acceleration response spectra. \myremove{w}\myrevision{W}e also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies. \myremove{Similarly} \myrevision{Also}, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.

We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake \myrevision{(one earthquake and three velocity models)} in a particular region. In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, \myrevision{metrics,} etcetera) in order to arrive to a sufficiently sound and robust decision tree. The procedural steps laid out here, nonetheless, remain\myremove{s} valid. We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.
